Estimation of flow trajectories in a multi-lines network

### 2023-01-14 paper_complex_networks/paper_complex_networks_v1.tex
update article

>>>>

\subsection{Statement of the problem (version 2, added by GG)}
Let $f_{st}=n_{st}/n_{\bullet\bullet}$ be the proportion of $st$-trips (empirical distribution) and let $g_{st}$ be some prior guess on its shape (theoretical distribution). Also suppose a prior guess of row and margin constraints for $f_{st}$, respectively $\alpha_s$ and $\beta_t$, which has to verify 
\begin{equation}
	\label{init_margin}
	\frac{\alpha_i}{\alpha_j} = \frac{a_i}{a_j} \qquad \frac{\beta_i}{\beta_j} = \frac{b_i}{b_j}  \qquad \forall i, j \text{ which are not true junctions.} \\
\end{equation}
Setting these gives us
\begin{equation}
	\label{flow_sum}
	n_{\bullet \bullet} = \frac{a_i}{\alpha_i} = \frac{b_i}{\beta_i} \qquad \forall i \text{ which are not true junctions.}  
\end{equation}
We are interested in finding
\begin{align}
	\label{constr_MaxEnt}
	\min_{\mathbf{f}\in\mathcal{F}} &\; \sum_{st}f_{st}\log \frac{f_{st}}{g_{st}}, \notag \\
	s.t. &\; \sum_t f_{st} = \alpha_s, \notag \\
	&\; \sum_s f_{st} = \beta_t.
\end{align}
The Lagragian is
\begin{equation}
	L = \sum_{st}f_{st}\log \frac{f_{st}}{g_{st}} - \sum_s \lambda_s (\alpha_s - \sum_t f_{st}) - \sum_t \mu_t (\beta_t - \sum_s f_{st}), \notag
\end{equation}
which gives, after deriving and setting to zero,
\begin{equation}
	\label{Sol}
	f_{st} = \phi_s \psi_t g_{st} \qquad \text{with } \phi_s := \exp(- 1 - \lambda_s) \text{, } \psi_t := \exp(- \mu_t).
\end{equation}
Using constraints in (\ref{constr_MaxEnt}), we find
\begin{equation}
	\label{Sol_LagMult}
	\phi_s = \frac{\alpha_s}{\sum_t \psi_t g_{st}}, \qquad \psi_t = \frac{\beta_t}{\sum_s \phi_s g_{st}}, 
\end{equation}
which gives an \emph{iterative fitting algorithm}. Starting with any $\psi^{(0)}_t > 0$, we do
\begin{equation}
	\label{Iterative fitting}
	\phi^{(\iota)}_s = \frac{\alpha_s}{\sum_t \psi^{(\iota)}_t g_{st}}, \qquad \psi^{(\iota + 1)}_t = \frac{\beta_t}{\sum_s \phi^{(\iota)}_s g_{st}}, 
\end{equation}

\subsection{Margin distributions update (added by GG)}

With equation (\ref{flow_sum}), we can obtain the flow associated to $f_{st}$
\begin{equation}
	\label{flow_from_distrib}
	n_{st} = n_{\bullet \bullet} f_{st}, 
\end{equation}
and with the edge-trip incidence matrix $\bm{\chi}$ (\ref{edgetrip}), we obtain the intra-line edge flows $\mathbf{Y}=(y_{ij})$ and the transfer edge flows $\mathbf{Z}=(z_{ij})$. \\

We now have to make sure constraints (\ref{entrer}) and (\ref{sortir}) are verified. The new margins distribution are set to 
\begin{align}
	\widetilde{\alpha}_s &= \frac{\min(0, a_s - z_{\bullet s})}{\sum_{s'} \min(0, a_{s'} - z_{\bullet {s'}})} \label{alpha_update} \\
	\widetilde{\beta}_t &= \frac{\min(0, b_t - z_{t \bullet})}{\sum_{t'} \min(0, b_{t'} - z_{{t'} \bullet})} \label{beta_update}
\end{align}

\subsection{Prior distribution update (added by GG)}

Transfer edges $(i, j)$ where $z_{i \bullet} > b_i$ or $z_{\bullet j} > a_j$ typically denote an overflowing tansfer flow regarding the $g_{st}$ prior distribution.  The latter must be adjusted in order to have $\mathbf{Z}$ comptible with constraints. For any edge $(i, j)$, let us compute the \emph{flow ratio} $r_{ij}$, with 
\begin{equation}
	\label{flow_ratio}
	r_{ij} = \max \left(1, \frac{z_{i \bullet}}{b_i}, \frac{z_{\bullet j}}{a_j} \right),
\end{equation}
where $r_{ij} > 1$ typically denotes an overflow through edge $(i, j)$. This flow ratio can be traced back to origin-destination couples $s, t$ in order to compute the \emph{orgin-destination flow ratio} $\bar{r}_{st}$ 
\begin{equation}
	\label{st_flow_ratio}
	\bar{r}_{st} = \max_{ij} \chi_{ij}^{st} r_{ij},
\end{equation}
where again $\bar{r}_{st} > 1$ denotes an overflow between $s$ and $t$. To ajdust the flow, we can devide the previous flow by this ratio
\begin{equation}
	\label{update_flow}
	\widetilde{n}_{st} =\frac{n_{st}}{\bar{r}_{st}}
\end{equation}
And the new prior distribution should follow 
\begin{equation}
	\label{update_distrib}
	\widetilde{g}_{st} = \frac{\left( \frac{\widetilde{n}_{st}}{\phi_s \psi_t} \right)}{\sum_{s',t'} \left( \frac{\widetilde{n}_{s',t'}}{\phi_{s'} \psi_{t'}} \right)}
\end{equation}

With the new prior distribution $\widetilde{g}_{st}$ and the new margin distributions $\widetilde{\alpha}_s$, $\widetilde{\beta}_t$, we can begin a new iteration step by going through 2.5 -2.7 again.
